---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


This is my 'Hello World' Mardown - coincidentally it is also the first Peer Assessment of the Reproducible Research course of Coursera's Data Science Specialization 

## Loading and preprocessing the data

```{r dataload, echo=T}
data <- read.table(unz("activity.zip", "activity.csv"), header = TRUE, sep=",")
data$date <- as.Date(data$date, format = "%Y-%m-%d")
```

## What is mean total number of steps taken per day?



```{r libraryload, message=F, warning=F, echo=T}
library(dplyr)
library(ggplot2)
data2 <- data[complete.cases(data),]
data2 <- group_by(data2, date)
dailysteps <- summarize(data2, total = sum(steps))
plot1 <- qplot(total, data = dailysteps) 
      + geom_histogram(fill = "steelblue")
      + labs(x = "Total number of steps taken each day", y = "Frequency", title = "Daily Total Steps Histogram")
plot1
summary(dailysteps$total)
plot1 + geom_vline(aes(xintercept = mean(dailysteps$total)), colour = "red", size = 1) 
      + geom_vline(aes(xintercept = median(dailysteps$total)), colour = "blue", size = 1, alpha = .5)
```

## What is the average daily activity pattern?

```{r dailyplot, message=F, warning=F, echo=T}
data3 <- group_by(data2,interval)
avgsteps <- summarize(data3, avg = mean(steps))
g <- ggplot(avgsteps,aes(x = interval, y = avg))
plot2 <- g + geom_line(size = 1, colour = "steelblue") 
          + labs(x = "Time interval [every 5 mins]", y = "Avg Steps", title = "Daily Avg of Steps taken by Interval")
plot2
max_step_interval <- avgsteps[avgsteps$avg==max(avgsteps$avg),]
max_step_interval
plot2 + geom_vline(aes(xintercept = max_step_interval$interval), colour = "red", size = 1, alpha = .2)
```

## Imputing missing values

```{r missingvalues, message=F, warning=F, echo=T}
missing <- count(data[complete.cases(data)==FALSE,])
missing
```
## Are there differences in activity patterns between weekdays and weekends?
